---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: true
---
Author: T. Aristanto
Date: 2/14/2015

## Loading and preprocessing the data
Load the activity.csv file from the workspace directory
No. 1
```{r load data, echo = TRUE}
actdata <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
Calculate the daily sum of steps. 
```{r daily steps info, echo = TRUE}
library(dplyr)
## Calculate Daily Sum
sum_daily <- actdata %>% group_by(date) %>% summarise(sumsteps = sum(steps, na.rm=TRUE))
sum_daily
```

Create the Histogram.
```{r create histogram 1, echo = TRUE}
## Create the Histogram
hist(as.numeric(sum_daily$sumsteps), breaks = 30, main = "Histogram of Daily Steps taken", xlab = "Steps taken", ylab ="Frequency")
```

Get the mean and median number.
```{r get the mean and median 1, echo = TRUE}
summary(sum_daily)
```

## What is the average daily activity pattern?
No.1 create the graph of the average daily pattern.
```{r daily pattern, cache = FALSE, echo = TRUE}
sum_daily2<- actdata %>% group_by(interval) %>% summarise(meansteps = mean(steps, na.rm=TRUE))
```

Plot the data as time interval 
```{r time interval 1, echo = TRUE}
library(ggplot2)
ggplot(sum_daily2, aes(as.numeric(interval), meansteps)) + geom_line() + scale_x_discrete(breaks=seq(0, 2400, by=100)) + ggtitle("Average steps every 5 minutes observations") + xlab("5 Minute Observation Interval (in minutes)") + ylab("Steps taken")
```

No.2 From The function below it can be seen that the max steps occurs at interval 835:
```{r daily max, echo = TRUE}
filter(sum_daily2, meansteps == max(meansteps))
```

## Imputing missing values
Replace the NA values with the 5 minutes interval average calculated above and then re calculate the daily sum
```{r imputing missing value, echo = TRUE}
library(dplyr)
actdata2 <- merge(actdata,sum_daily2, by.X="interval")
actdata3 <- actdata2 %>% mutate(goodsteps = ifelse(is.na(steps),meansteps,steps)) %>% select(goodsteps, date, interval) %>% arrange(date,interval) 
sum_daily3 <- actdata3 %>% group_by(date) %>% summarise(sumsteps = sum(goodsteps, na.rm=TRUE))
```

Recreate the Histogram from the new data
```{r create histogram 2, echo = TRUE}
hist(as.numeric(sum_daily3$sumsteps), breaks = 30, main = "Histogram of Daily Steps taken (NA cleaned)", xlab = "Steps taken", ylab ="Frequency")
```

Calculate the mean and median of the data
```{r get mean and median 2, echo = TRUE}
summary(sum_daily3)
```

The Graph shows that less frquency of 0 steps
Also the mean and the median are closer (equal) to each other. 


## Are there differences in activity patterns between weekdays and weekends?
Add column to identify weekday and weekend the calculate the average interval.
```{r Weekdays vs weekend, echo = TRUE}
actdata4 <- actdata3 %>% mutate(weektype = ifelse(weekdays(as.Date(date)) %in% c("Saturday","Sunday"),"weekend","weekday"))
meanweek <- actdata4  %>% group_by(weektype,interval) %>% summarise(meansteps = mean(goodsteps, na.rm=TRUE))
```

Plot the graph for weekday and weekend
```{r time interval 2, echo = TRUE}
library(ggplot2)
ggplot(meanweek, aes(x=interval, y=meansteps)) + geom_line() + scale_x_discrete(breaks=seq(0, 2400, by=100)) + facet_grid(weektype ~ .) + ggtitle("Weekday vs Weekend Activities") + xlab("5 Minute Observation Interval (in minutes)") + ylab("Steps taken")
```